---
title: "Willow_statistics"
author: "Nora Bull"
date: "25 8 2021"
output: html_document
---



### Willow statistics


# Imports
```{r, eval=FALSE}
library(dplyr)
library(plyr)
```

# Hypothesis test

### Calculation of p-values for all GDRs

Procedure to find number of samples needed for 90% significance level:
  The variable `num_perm_samples` was edited and tested for multiple values, until `high_q` had a value of ~0.9. 

```{r}
prob = 0.05
var_p = prob * (1 - prob)
num_perm_samples = 100    # 70 = empirical p val = 0.1
var_p_n = var_p / num_perm_samples
sd_p_n = sqrt(var_p_n)

# 0.025 quantile
low_q = max(0, (prob-1.96*sd_p_n))

# 0.0975 quantile of true empirical p-val
high_q = prob + 1.96 * sd_p_n

print(low_q)
print(high_q)
```

**Interpretation:** 
True empirical p-value for 100 samples is between 0.009 and 0.091. 

For 1000 samples, it is only dropped to 0.036 - 0.0635, so around 3% more 
confidence. It is not worth the computational time here. 




Cumulative distribution function

- Calculate p-value for each distribution.
- Create function 


## Function to perform p-val caluclation for one file
```{r}

# Dummy
#GDRnullFile <- superFiles[3]
#GDRnullFile <- "C:/Users/norab/Master/Data/GDRnull/refined_values/super/GDRnullSuper_ENSG00000178814___OPLA.csv"
#GDRnullFile <- "C:/Users/norab/Master/Data/GDRnull/refined_values/super/GDRnullSuper_ENSG00000134121___NCHL1.csv"
#GDRnullFile <- "C:/Users/norab/Master/Data/GDRnull/refined_values/super/GDRnullSuper_ENSG00000284869___SELB.csv"


#pvalSaveFile <- file_super
#GDRs <- GDRsuper

calcAndSave_pval <- function(GDRnullFile, GDRs, pvalSaveFile) {
  
  # Read data
  GDRnullData <- read.csv(GDRnullFile) # Swap to file
  gene.name <- gsub("^.*_EN", "EN", GDRnullFile)
  gene.name <- gsub(".csv$", "", gene.name)
  cat(gene.name)
  cat('\n')
  GDRval <- GDRs$GDR[GDRs$gene == gene.name]

  P = ecdf(GDRnullData$SDRnull)    # The function giving empirical CDF of GDRnull_values
  #plot(P)
  p_val = P(GDRval)                # Returns empirical CDF at GDRtrue, ie. calculated GDR for populations
                                   # Save p val for next step - multiple testing correcting
  #plot(P)                         # Plot the shit (dont do for all, make an example or two...
  #cat(p_val)
  # Save p val to file
  cat(c(gene.name, p_val), file = pvalSaveFile, sep = ",", append = TRUE)
  cat('\n', file = pvalSaveFile, append=TRUE)
}
```


## Run GDR calulcation for all trees
```{r}

# Load calculated GDRs
GDRsuper <- read.csv("C:/Users/norab/Master/data/GDR/GDRsuper_all.csv", header = TRUE)
GDRsub <- read.csv("C:/Users/norab/Master/data/GDR/CALC_P_GDR_SUB/GDRsub_all.csv", header = TRUE)

# Create file to save p values:
file_super <- "C:/Users/norab/Master/data/GDRnull/refined_values/GDRnull_pval_super.csv"
file.create(file_super)
#file_sub <- "C:/Users/norab/Master/data/GDRnull/refined_values/GDRnull_pval_sub.csv"
file_sub <- "C:/Users/norab/Master/data/GDR/CALC_P_GDR_SUB/GDRnull_pval_sub.csv"
file.create(file_sub)

# Create filelist to process
superFiles <- list.files(path="C:/Users/norab/Master/data/GDRnull/refined_values/super", full.names=TRUE, recursive=FALSE)
subFiles <- list.files(path="C:/Users/norab/Master/data/GDRnull/refined_values/sub", full.names=TRUE, recursive=FALSE)
# Action

super <- sapply(superFiles, calcAndSave_pval, GDRsuper, file_super)
sub <- sapply(subFiles, calcAndSave_pval, GDRsub, file_sub)


```

# Multiple testing correcting
```{r}

# Read data
p_vals_super <- read.csv("C:/Users/norab/Master/data/GDRnull/refined_values/GDRnull_pval_super.csv", sep=",", col.names = c("gene", "pval"))
p_vals_sub <- read.csv("C:/Users/norab/Master/data/GDRnull/refined_values/GDRnull_pval_sub.csv", sep=",", col.names = c("gene", "pval"))

# adjust
p_vals_super$pval_adj <- p.adjust(p_vals_super$pval)
p_vals_sub$pval_adj <- p.adjust(p_vals_sub$pval)

# Add GDR
SUPERdata <- inner_join(p_vals_super, GDRsuper, by = "gene")
SUBdata <- inner_join(p_vals_sub, GDRsub, by = "gene")

throw_genes_super <- p_vals_super$gene[p_vals_super$pval_adj > 0.001 ] 
keep_genes_super <- SUPERdata[SUPERdata$pval_adj < 0.001,]
keep_genes_sub <- SUBdata[SUBdata$pval_adj < 0.001,]

keep_genes_super_GDR_50 <- SUPERdata[SUPERdata$GDR < 0.62,]
```

# Save gene list
```{r}
genes = keep_genes_super$gene
genes50 = keep_genes_super_GDR_50$gene
genes2 <- sapply(genes, function (x) gsub("^ENS.*___", "", x))
genes3 <-  sapply(genes, function (x) gsub("___.*$", "", x))
genes50lowest <- sapply(genes50, function (x) gsub("___.*$", "", x))

# Save
lapply(genes2, write, "C:/Users/norab/Master/data/GDR/GDR1005_significantGenes_GDRlimit0.8.csv", append=TRUE)
lapply(genes, write, "C:/Users/norab/Master/data/GDR/GDR1005_significantGenesFullname.csv", append=TRUE)
lapply(genes3, write, "C:/Users/norab/Master/data/GDR/GDR1005_significantGenesFullname_ENS.csv", append=TRUE)
lapply(genes50lowest, write, "C:/Users/norab/Master/data/GDR/50LowestSignificantGenes_ENS.csv", append=TRUE)


lapply(SUPERdata, write, "C:/Users/norab/Master/data/GDR/GDR1005_significantGenesSUPER.csv", append=TRUE)
write.table(SUBdata, file = "C:/Users/norab/Master/data/GDR/GDR1005_significantGenesSUB.csv", quote = FALSE, sep = ",", row.names = FALSE)



```

#Plot null dist to check
```{r}
GDRnullFile <- "C:/Users/norab/Master/thesis_data/GDRnull/super/SDRnull_ENSG00000134121___NCHL1.csv"
GDRnullFile <- "C:/Users/norab/Master/thesis_data/result_data/GDRnull/super/SDRnull_ENSG00000000971___CFAH.csv"

GDRsuper <- read.csv("C:/Users/norab/Master/thesis_data/result_data/GDR/GDRsuper_all.csv", header = TRUE)


GDRnullData <- read.csv(GDRnullFile) # Swap to file
gene.name <- gsub("^.*_EN", "EN", GDRnullFile)
gene.name <- gsub(".csv$", "", gene.name)
GDRval <- GDRsuper$GDR[GDRsuper$gene == gene.name]
P = ecdf(GDRnullData$SDRnull)    # The function giving empirical CDF of GDRnull_values
p_val = P(GDRval)                # Returns empirical CDF at GDRtrue, ie. calculated GDR for populations
                                   # Save p val for next step - multiple testing correcting
plot(P, 
     main = "GDRnull: NCHL1         GDR = 0.9757", 
     xlab = "GDR", 
     ylab = "cumulative %",
     cex.lab = 1.5,    # X-axis and Y-axis labels size
     cex. = "sans-serif")  

```


```{r}
GDRnullFile <- "C:/Users/norab/Master/Data/GDRnull/refined_values/sub/GDRnullSub_ENSG00000168653___NDUS5.csv"
GDRsub <- read.csv("C:/Users/norab/Master/data/GDR/GDRsub_all.csv", header = TRUE)


GDRnullData <- read.csv(GDRnullFile) # Swap to file
gene.name <- gsub("^.*_EN", "EN", GDRnullFile)
gene.name <- gsub(".csv$", "", gene.name)
GDRval <- GDRsub$GDR[GDRsub$gene == gene.name]
P = ecdf(GDRnullData$GDRnull)    # The function giving empirical CDF of GDRnull_values
p_val = P(GDRval)                # Returns empirical CDF at GDRtrue, ie. calculated GDR for populations
                                   # Save p val for next step - multiple testing correcting
plot(P, main = "GDRnull: NDUS5         GDR = 0.953", xlab = "GDR", ylab = "cumulative %")  

```

### Extract relevant GDRnull values
```{r}

# Read all GDRnull data
GDRnull_all_super <- read.csv("C:/Users/norab/Master/data/GDRnull/all/GDRnullTotalSuper_22.07.21.csv")
GDRnull_all_sub <- read.csv("C:/Users/norab/Master/data/GDRnull/all/GDRnullTotalSub_22.07.21.csv")

# Read list of relevant genes
GDRnull_genes <- read.csv("C:/Users/norab/Master/data/GDRnull/other/GDRnull_gene_selection.csv")
colnames(GDRnull_genes) <- c("gene")

# Extract GDRnull data for selected genes

# Info
gene_nr <- table(GDRnull_all_super['gene'])
max(gene_nr)

# Get generated GDRnull for selected genes only
GDRnull_selection_super <- dplyr::filter(GDRnull_all_super, GDRnull_all_super$gene %in% GDRnull_genes$gene)
GDRnull_selection_sub <- dplyr::filter(GDRnull_all_sub, GDRnull_all_sub$gene %in% GDRnull_genes$gene)

# Next: filter genes with more than 
CFAH_example <- filter(GDRnull_selection_super, gene == "ENSG00000000971___CFAH")

# Load GDRs for population groups
GDRsuper <- read.csv("C:/Users/norab/Master/data/GDR/GDRsuper_all.csv", header = TRUE)
GDRsub <- read.csv("C:/Users/norab/Master/data/GDR/GDRsub_all.csv", header = TRUE)



GDR_CFAH <- filter(GDRsuper, gene == "ENSG00000000971___CFAH")
GDR_RBL1 <- sum(filter(keep_genes_super, gene == "ENSG00000080839___RBL1"))



```



```{r}


any(SUPERdata$gene == "ENSG00000110700___RS13")
SUPERdata[SUPERdata$gene ==  "ENSG00000110700___RS13",] 
SUPERdata[SUPERdata$gene ==  "ENSG00000080839___RBL1",]
SUBdata[SUBdata$gene ==  "ENSG00000080839___RBL1",]
GDRsuper[GDRsuper$gene ==  "ENSG00000080839___RBL1",]
p_vals_super[p_vals_super$gene ==  "ENSG00000080839___RBL1",]


```



# Plot null distribution

```{r}
install.packages("ggplot")
library(ggplot2)
# Filled Density Plot
ggplot(GDRnull, aes(x = GDRnull)) + 
       geom_histogram(color = "red", # Curve color
                    fill = "purple",  # Area color
                    alpha = 0.5)   # Area transparency
```


# Willow statistics simulation

```{r}
GDRnullFile<- "C:/Users/norab/Master/thesis_data/simulation/simData/randSimG9_P5.csv"

calcAndSave_simNull_pval <- function(GDRnullFile, pvalSaveFile) {
  
  # Read data
  GDRnullData <- read.csv(GDRnullFile)
  col = colnames(GDRnullData)
  data = unlist(strsplit(col, '_'))
  #G = strtoi(gsub("G.", "", data[1]))
  G = gsub("G.", "", data[1])
  perm = gsub("P.", "", data[2])
  GDR = as.double(gsub("GDR.", "", data[3]))
  uniq_pvals = length(unique(GDRnullData[,1]))
  
  
  P = ecdf(GDRnullData[,1])    # The function giving empirical CDF of GDRnull_values
  p_val = P(GDR)               # Returns empirical CDF at GDRtrue, ie. calculated GDR for populations
                               # Save p val for next step - multiple testing correcting
  plot(P)                      # Plot the shit (dont do for all, make an example or two...
  # Save p val to file
  #GDR <- as.character(GDR)
  cat(c(G, perm, GDR, uniq_pvals, p_val), file = pvalSaveFile, sep = ",", append = TRUE)
  cat('\n', file = pvalSaveFile, append=TRUE)
}


```


# Run the shit
```{r}

# Create file to save p values:
simNull_pvals_file <- "C:/Users/norab/Master/data/simulation/simNull_pvals.csv"
file.create(save_file)
# Write header to file
cat(c("group_size", "perm", "GDR", "n_uniq_pvals", "pval"), file = simNull_pvals_file, append=TRUE)

# Create filelist to process
simNullFiles <- list.files(path="C:/Users/norab/Master/thesis_data/simulation/simData/", full.names=TRUE, recursive=FALSE)

# Action
test <- sapply(simNullFiles, calcAndSave_simNull_pval, simNull_pvals_file)

```

# Multiple testing correcting
```{r}

# Read data
p_vals <- read.csv("C:/Users/norab/Master/data/simulation/simNull_pvals.csv", sep=",")

# adjust
p_vals$pval_adj <- p.adjust(p_vals$pval)


# Save to file
write.csv(p_vals, "C:/Users/norab/Master/data/simulation/simNull_pvals.csv", row.names = FALSE, quote = FALSE)


```


```{r}
prob = 0.05
var_p = prob * (1 - prob)
num_perm_samples = 700      # 70 = empirical p val = 0.1
var_p_n = var_p / num_perm_samples
sd_p_n = sqrt(var_p_n)

# 0.0975 quantile of true empirical p-val
high_q = prob + 1.96 * sd_p_n

print(high_q)
```