---
title: "GDRsimulationStats"
author: "Nora Bull"
date: "18 11 2021"
output: html_document
---


# Willow statistics simulation

Notes: 
- Remove all "dummy's"

## Function to calculate and save p-values for simulated GDR
```{r}
GDRsim_pval_calcSave <- function(GDRnull_file, pvalSaveFile) {
  
  "
  Input:
    GDRnull_file: path to file with random GDR values 
      Format: CSV
      
    pval_save_file: desired filepath to store values
      Format: CSV
  "
  
  # Read data
  GDRnull_data <- read.csv(GDRnull_file)
  col = colnames(GDRnull_data)
  data = unlist(strsplit(col, '_'))
  G = gsub("G", "", data[1])
  perm = gsub("P", "", data[2])
  GDR = as.double(gsub("GDR", "", data[3]))
  uniq_pvals = length(unique(GDRnull_data[,1]))
  
  # Create empirical cumulative distribution function (ecdf)
  P = ecdf(GDRnull_data[,1])
  
  # Calculate p value for GDR
  p_val = P(GDR)

  # Save p val to file
  #GDR <- as.character(GDR)
  cat(c(G, perm, GDR, uniq_pvals, p_val), file = pvalSaveFile, sep = ",", append = TRUE)
  cat('\n', file = pvalSaveFile, append=TRUE)
}


```
_
# Run p-value calculation
```{r}

# Create file to save p-values:
save_file <- "C:/Users/norab/Master/thesis_data/simulation/simNull_pvals_24.11.21.csv"
file.create(save_file)
cat(c("group_size,permutations,GDR,n_uniq_pvals,pval\n"), file = save_file, append=TRUE)

# Create list of files to process (each containing random GDRs for one simulated GDR values)
simNull_files <- list.files(path="C:/Users/norab/Master/thesis_data/simulation/simData/", full.names=TRUE, recursive=FALSE)

# Apply GDRsim_pval_calcSave function to all simNull_files
run <- sapply(simNull_files, GDRsim_pval_calcSave, save_file) 

```

# Multiple testing correction
```{r}

# Read data
all_data <- read.csv(save_file, sep=",")

# adjust
all_data$pval_adj_holm <- p.adjust(all_data$pval)
all_data$pval_adj_bonferroni <- p.adjust(all_data$pval, method = "bonferroni")
all_data$hommel <- p.adjust(all_data$pval, method = "hommel")

# Save
save_to <- 'C:/Users/norab/Master/thesis_data/simulation/simNull_alldata_24.11.21.csv'
write.csv(all_data, save_to, row.names = FALSE)
```


# Plot CDF for four simulated instances

Retrieve ecdf and p_val for 4 files:

The 4 scenarios of simulated GDR value distributions are plotted: 
G = number of group samples
P = number of permuted samples

  1. Low G, low P
  2. Low G, high P
  3. High G, low P
  4. High G, low P

Low = ~20 % permuted samples
High = ~80 % permuted samples

## Function to plot ecdf and
```{r}

ecdf_pval <- function(GDRnullFile, title) {
  
  # Read data
  GDRnullData <- read.csv(GDRnullFile)
  col = colnames(GDRnullData)
  data = unlist(strsplit(col, '_'))
  G = gsub("G", "", data[1])
  perm = gsub("P", "", data[2])
  GDR = as.double(gsub("GDR", "", data[3]))
  
  # Make ecfd
  P = ecdf(GDRnullData[,1])
  # calc p-value
  p_val = P(GDR)
  
  # Plot
  title = paste0("simGDR: ", title)
  plot(P, main = title, xlim=c(0,1), ylab="Probability", xlab="GDR")          
  
  abline(v=GDR, col="red", lwd = 2, lty = 2)
  abline(h=p_val, col="purple", lwd = 2, lty = 2)
  GDRlab = paste("GDR = ",round(GDR, 2))
  p_valLab = paste("p-val = ", p_val)
  legend(x=0, y=1,c(GDRlab, p_valLab), col=c("red", "purple"), lty=2, cex=0.8)
  
  print(p_val)
  return(GDR)}

```

## Plot
```{r}
# Files to plot
randSimG20_P4 <- "C:/Users/norab/Master/thesis_data/simulation/simData/randSimG20_P4.csv"
randSimG20_P16 <- "C:/Users/norab/Master/thesis_data/simulation/simData/randSimG20_P16.csv"
randSimG190_P38 <- "C:/Users/norab/Master/thesis_data/simulation/simData/randSimG190_P38.csv"
randSimG190_P152 <- "C:/Users/norab/Master/thesis_data/simulation/simData/randSimG190_P152.csv"

# Plot
ecdf_pval(randSimG20_P4, title = "G20 / P4")
ecdf_pval(randSimG20_P16, title = "G20 / P16")
ecdf_pval(randSimG190_P38, title = "G190 / P38")
ecdf_pval(randSimG190_P152, title = "G190 / P152")

```


