---
title: "GOenrichment"
author: "Nora Bull"
date: "25 5 2021"
output: html_document
---



```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("topGO")
BiocManager::install("biomaRt")
install.packages("Rgraphviz")
```

```{r}
library(topGO)
library(GO.db)
library(biomaRt)

```

## 1: Preparing the data in the required format

Get list of genes of interest and list of all genes (gene universe). 
Gene universe is all genes in Reactome database that was used in the clustering
process. 

*Identifiers:* Ensemble ID  
```{r}
# Gene Universe (all genes in my experiment)
geneUni = read.table('C:/Users/norab/MasterDisaster/Data/go_enrichment/geneUniverse.csv', header=TRUE, sep='')
geneUniEns <- gsub("___.*$", '', geneUni$gene)
geneUniName <- gsub("^.*___", '', geneUni$gene)


superSDR.07 = read.table('C:/Users/norab/MasterDisaster/Data/go_enrichment/superSDRgenes07.csv', header=TRUE, sep='')
superSDR.07 = gsub("___.*$", '', superSDR.07$gene)

superSDR.06 = read.table('C:/Users/norab/MasterDisaster/Data/go_enrichment/superSDRgenes07.csv', header= 'TRUE', super='___')
superSDR.06 = as.character(superSDR.07)

superSDR.05 = read.table('C:/Users/norab/MasterDisaster/Data/go_enrichment/superSDRgenes07.csv', header= 'TRUE', super='___')
superSDR.05 = as.character(superSDR.07)

```

Inspection 
```{r}
length(geneUniEns)
```
```{r}
head(geneUniEns)
```
```{r}
length(superSDR.07)
```
```{r}
head(superSDR.07)
```



## 2: Create GO annotation
Get GO terms and GO mappings that associate each gene 
in this study with one or more GO terms
```{r}
# create GO db for genes to be used using biomaRt
db= useMart('ENSEMBL_MART_ENSEMBL',dataset='hsapiens_gene_ensembl', host="www.ensembl.org")

# Create the ENSEMBL to GO ID mapping. 
# Bea aware to select the correct attributes! 
go_ids= getBM(attributes=c('go_id', 'ensembl_gene_id'), filters='ensembl_gene_id', values=geneUniEns, mart=db)

```
List of valid attributes / filters / Marts
```{r}
listAttributes(db)
listMarts()
listFilters(db)
```


```{r}
# build the gene 2 GO annotation list (needed to create topGO object)
gene_2_GO=unstack(go_ids[,c(1,2)])
 
# remove any candidate genes without GO annotation
keep = superSDR.07 %in% go_ids[,2]
keep =which(keep==TRUE)
candidate_list=superSDR.07[keep]
 
# make named factor showing which genes are of interest
geneList=factor(as.integer(geneUniEns %in%candidate_list))
names(geneList)= geneUniEns

```



## 3: Make topGO data object

This data object will contain all gene identifiers and their scores (hmmm....),
the GO annotations, the GO hierarchical structure and all other info
needed to perform the desired enerichment analysis.

Need: 
  1. Ontology: character string specifying the ontology of itnerest (BP, MF, CC)
    BP = biological process
    CC = cellular component
    MF = molecular function
  2. allGenes: named vector of type numeric or factor. The names attribute contains the gene
  identifiers. The genes listed in this object define the gene universe.
  3. nodeSize: Integer >= 1. This parameter is used to prune the GO hierarchy from
  the terms which have less than nodeSize annotated genes (after true path rule
  is applied).
  4. annot - annotationFun: function which maps gene identifiers to GO terms. There are a couple
  of annotation function included in the package trying to address the users needs. 
  The annotations functions take 3 arguments. One of those arguments is specifying where the mappings can be found, and needs to be provided by the user. 
  annFUNN.gene2GO - this function is used when the annotations are provided as a 
  gene-to-GOs mapping. 
  
*My notes*:

*Difference from topGO tutorial:*
The argument "geneSel" = topDiffGenes is not present in my analysis. This is meant to describe the top 50 genes with a raw p value less than 0.01 out of the provided genes (323 in their example).
I dont have p-values like this, but I may when I make a null distrubutuin for SDR?

*Parameters:*



```{r}
GOdata=new('topGOdata', 
           description = "first test session",
           ontology='BP',
           nodeSize = 10, # 
           allGenes = geneList,  
           annot = annFUN.gene2GO, 
           gene2GO = gene_2_GO)

GOdata
```



# 4: Performing the enrichment test
Requirement: the an object of class topGOdata, 
Two types of test statistics is used: 
1. Fisher's exact test (based on gene counts)
2. Kolmagorov-Smirnov like test - computes enrichemnt based on gene scores
(dont think I can use this...)

Can use both test when each gene has a score represetning how differentially expressed a gene is.
It should thenn be categorized into differentially expressed genes (or not) 
by  the means of topDiffGenes functions (which is not possible in my case). 

In my case: will *NOT* perform an enrichment analysis, but an Over-representation analysis. 
This only requires a list of identifiers and analyses whether the list I supply is 
significant associated with a particular pathway or set of pathways.
This means it will be localised to certain pathways or classifications
instead of random scattered thorughout the whole set of possible pathways. 
The list of identifiers I have is a subset of all the metabolites I have/have measured. 
It relies on me having selected the sub-set appropriately.
NB: Check if all genes are treated as equally important by the test.
It should not be the case, as SDR values are different and I think it could be
meaningful to treatthem in a sorted manner. 
Find test for this. 



```{r}
# define test using the classic algorithm with fisher (refer to [1] if you want to understand how the different algorithms work)
classic_fisher_result=runTest(GOdata, algorithm='classic', statistic='fisher')

```

Comment on p-values: these are unadjusted for multiple testing. 
Adjusting p-values may be a good idea, but can also be misleading. Be aware of this!


Over-representation enrichment
When algorithm = classic, the GO hierarchy isn't taken into account, so each GO term is tested independently (over-representation enrichment). The limitation of this is that all genes annotated to a GO term will be automatically annotated to its parents as well, therefore a GO term might look enriched just because its children are enriched.
Thus, it is important that GO hierarchy is taken into account (conditional enrichment) to avoid redundancy.

```{r}
# define test using the weight01 algorithm (default) with fisher
weight_fisher_result=runTest(GOdata, algorithm='weight01', statistic='fisher') 
 
```


## 5: Correcting for multiple testing
If p-value interpretation is very important. But again, be aware of
interpretations as it may be misleading. 

```{r}
#performing BH correction on our p values
p.adj=round(p.adjust(all_res$weightFisher,method="BH"),digits = 4)
 
# create the file with all the statistics from GO analysis
all_res_final=cbind(all_res,p.adj)
all_res_final=all_res_final[order(all_res_final$p.adj),]
 
#get list of significant GO before multiple testing correction
results.table.p= all_res_final[which(all_res_final$weightFisher<=0.001),]
 
#get list of significant GO after multiple testing correction
results.table.bh=all_res_final[which(all_res_final$p.adj<=0.05),]
 
#save first top 50 ontolgies sorted by adjusted pvalues
write.table(all_res_final[1:50,],"summary_topGO_analysis.csv",sep=",",quote=FALSE,row.names=FALSE)
 
# PLOT the GO hierarchy plot: the enriched GO terms are colored in yellow/red according to significance level
 
pdf(file='topGOPlot_fullnames.pdf', height=12, width=12, paper='special', pointsize=18)
showSigOfNodes(GOdata, score(weight_fisher_result), useInfo = "none", sigForAll=FALSE, firstSigNodes=2,.NO.CHAR=50)
dev.off()
```

## Analysis of the result: 

GenTable is an easy to use function for analysing the most significant GO terms and the corresponding p-values (these
are now somehting I have, from what I understand). 

```{r}
# generate a table of results: we can use the GenTable function to generate a summary table with the results from tests applied to the topGOdata object.
allGO=usedGO(GOdata)
all_res=GenTable(GOdata, weightFisher=weight_fisher_result, orderBy='weightFisher', topNodes=length(allGO))
```


